---
title: "ABC7 Chicago Neighborhood Safety Tracker"
# author: "Maggie Green"
# date: Sys.time()
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(tidyverse)
library(leaflet)
library(leaflet.extras)
library(leaflet.providers)
library(htmlwidgets)
library(htmltools)
library(sf)

```

<link href="https://fonts.googleapis.com/css2?family=Roboto:wght@100;300;400;500;700;900&display=swap" rel="stylesheet">

![<popupchatter>Updated `r format(Sys.time(), '%B %d, %Y')` by the ABC7
Data
Team</popupchatter>](https://github.com/abcotvdata/safetytracker_chicago/raw/main/docs/banner_chicago.jpg)

```{css, echo=FALSE}

h1.title {
  font-family: roboto;
  color: transparent;
  font-weight: 700;
  text-align: left;
  font-size: 12px;
  padding: 0px;
}

.date {
  font-family: roboto;
  color: black;
  font-weight: 300;
  text-align: left;
  font-size: 10px;
    padding: 5px;
}

select {
  margin: 0px;
  width: 180px;
  color: #00318b;
  padding: 5px 35px 5px 5px;
    font-family: roboto;
  font-size: 18px;
  font-weight: 900;
  border: 0px;
  height: 34px;
  -webkit-appearance: none;
  -moz-appearance: none;
  appearance: none;
  background: url(https://upload.wikimedia.org/wikipedia/commons/9/9d/Arrow-down.svg) 96% / 15% no-repeat #f2f2f2;
}

h1 {
  font-family: roboto;
  color: black;
  font-weight: bolder;
  text-align: left;
  font-size: 36px;
  margin-top: 0;
  margin-bottom: 0;
}

h2 {
  font-family: roboto;
  font-weight: 500;
  color: black;
  text-align: center;
  font-size: 18px;
  margin-top: 0;
  margin-bottom: 2;
}

bignumber {
  font-family: roboto;
  color: #00318b;
  font-weight: 900;
  text-align: center;
  font-size: 60px;
    line-height: 65px;
  height: 65px;
    margin-top: 0;
  margin-bottom: 0;
}

subhead {
  font-family: roboto;
  color: black;
  font-weight: 700;
  text-align: left;
  font-size: 20px;
    padding: 0px;
}

body {
  color: black;
  font-family: roboto;
  font-weight: 400;
  font-size: 18px;
}

popuptitle {
  color: #00318b;
  font-family: roboto;
  font-weight: 700;
  font-size: 15px;
  text-align: left;
}

popupchatter {
  font-family: roboto;
  color: black;
  font-weight: 300;
  text-align: left;
  font-size: 12px;
  margin-top: 0;
  margin-bottom: 2;
}

h8 {
  color: #00318b;
  font-family: roboto;
  font-weight: 900;
  font-size: 18px;
}

table {
  font-family: roboto;
  width: 100%;
}

tr {
  border-bottom: thin solid #99a0a5;
}
  
td {
  text-align: right;
  padding: 1px;
}

th {
  text-align: right;
  padding: 1px;
}
  
   * {
      box-sizing: border-box;
   }
   .card {
      color: white;
      float: left;
      width: calc(25% - 10px);
      padding: 5px;
      border-radius: 10px;
      margin-left: 3px;
      margin-right: 3px;
      margin-top: 3px;
      margin-bottom: 3px;
      height: 100%;
   }
   .card p {
   font-family: roboto;
   text-align: center;
   font-size: 14px;
  margin-bottom: 0;
   }
   .cardContainer:after {
      content: "";
      display: table;
      clear: both;
   }
   @media screen and (max-width: 650px) {
      .card {
         width: 100%;
      }
      h1.title {
        font-size: 22px;
      }
   }
```

```{r population, include=FALSE}
# set value of chicago_population
chicago_population <- 2665064
```

```{r crimedata, include=FALSE}
# Source data pre-processed separately at process_chicago_crime.R
```

<h2>

ABC7 is tracking crime and safety across the city of Chicago and in your
neighborhood.

<br>You can choose which crime to examine:
<select onchange="window.location=this.value">
<option value="https://abcotvdata.github.io/safetytracker_chicago/Chicago_Safety_Tracker_Transit.html">CTA Violent Crimes</option>
<option value="https://abcotvdata.github.io/safetytracker_chicago/Chicago_Safety_Tracker_Robberies.html">Robberies</option>
<option value="https://abcotvdata.github.io/safetytracker_chicago/Chicago_Safety_Tracker_Batteries.html">Batteries</option>
<option value="https://abcotvdata.github.io/safetytracker_chicago/Chicago_Safety_Tracker_VehicleThefts.html">Vehicle
Thefts</option>
<option value="https://abcotvdata.github.io/safetytracker_chicago/Chicago_Safety_Tracker_Burglaries.html">Burglaries</option>
<option value="https://abcotvdata.github.io/safetytracker_chicago/Chicago_Safety_Tracker.html">Homicides</option>
<option value="https://abcotvdata.github.io/safetytracker_chicago/Chicago_Safety_Tracker_Robberies.html">Robberies</option>
<option value="https://abcotvdata.github.io/safetytracker_chicago/Chicago_Safety_Tracker_SexualAssaults.html">Sexual
Assaults</option>
<option value="https://abcotvdata.github.io/safetytracker_chicago/Chicago_Safety_Tracker_Thefts.html">Thefts</option>
<option value="https://abcotvdata.github.io/safetytracker_chicago/Chicago_Safety_Tracker_CTA.html">CTA
Crime</option> </select>

</h2>

<div class="cardContainer">

<div class="card" style="background-color:#99a0a5;">

<p>

CTA Violent Crime<br>
<bignumber>`r prettyNum(violent_transit$last12mos, big.mark=",")`<br></bignumber>

<p>Last 12 months<br>Through `r format(asofdate, "%B %d")`</p>

</div>

<div class="card" style="background-color:#99a0a5;">

<p>

Average Violent Crime on CTA<br>
<bignumber>`r prettyNum(round(violent_transit$avg_prior3years,0), big.mark=",")`<br></bignumber>

<p>Yearly average<br>2021 to 2023</p>

</div>

<div class="card" style="background-color:#99a0a5;">

<p>

Violent Crime Rate<br>
<bignumber>`r round(violent_transit$rate_last12,0)`<br></bignumber>

<p>Per 1,000,000 CTA riders<br>Last 12 months</p>

</div>

<div class="card" style="background-color:#99a0a5;">

<p>

Average Violent Crime Rate<br>
<bignumber>`r round(violent_transit$rate_prior3years,0)`<br></bignumber>

<p>Per 1,000,000 CTA riders<br>2021 to 2023</p>

</div>

</div>

<br> Violent crimes on CTA transit over the last 12 months are
<h8>`r ifelse(violent_transit$inc_prior3yearavgtolast12>=0," up"," down")`
`r ifelse(violent_transit$inc_prior3yearavgtolast12>=0,violent_transit$inc_prior3yearavgtolast12,substring(violent_transit$inc_prior3yearavgtolast12,2))`%</h8>
compared to the annual average over the last three years, according to
Chicago Police Department data through `r format(asofdate, "%B %d")`. However, more people are riding the CTA than in recent years, meaning that the rate of violent crimes per rider is much lower.

The city averaged <h8>`r round(violent_transit$last12mos/52,0)`</h8>
violent crimes on CTA trains and buses each week over the last 12
months. In 2019, that number was
<h8>`r round(violent_transit$total19/52,0)`</h8> a week.

<subhead>A closer look at Violent Crimes on CTA, neighborhood by
neighborhood</subhead>

The risk is not the same neighborhood to neighborhood. ABC7's data team
looked at the Chicago Police Department's data by neighborhood from 2019
through `r format(asofdate, "%B %d, %Y")`.

The map color codes each neighborhood by the violent crime rate on the
CTA over the last 12 months. The three darker blues represent
neighborhoods with violent crime rates that are higher than the citywide
rate. You can also click the box in the bottom right corner to see
neighborhoods by the number of violent crimes on the CTA.

Click on any neighborhood on the map to see detailed figures, rates and
trends. You can also search for a street name, place, landmark or zip
code.

<br>

```{r cars, echo=FALSE,out.width='100%',out.height='600'}
# CTA MAP

# Set bins for numbers of crimes for CTA map
ctabins1 <- c(0,
                20,
                50,
                100,
                500,
                1500)
ctabins2 <- c(0,
                10,
                25,
                50,
                100,
                200)
ctapal1 <- colorBin(c("#aecfe6",
                             "#2787d6",
                             "#0058f6",
                             "#003ca6",
                             "#00215c"), cta_area$rate_last12, bins = ctabins1,na.color = "#51585f")

ctapal2 <- colorBin(c("#aecfe6",
                             "#2787d6",
                             "#0058f6",
                             "#003ca6",
                             "#00215c"), cta_area$last12mos, bins = ctabins2,na.color = "#51585f")


# Create quick labels for CTA map
ctalabel <- paste(sep="",
                     "<popuptitle>",
                     cta_area$last12mos,
                     " Violent Crimes on CTA </popuptitle><br>",
                     " in the last 12 months in <b>",
                     cta_area$community,
                     "</b>, home to an estimated ",
                     prettyNum(cta_area$population, big.mark=","),
                     " people.
<br>
<table>      
      <tr>
				<th></th>
				<th>Total</th>
				<th>Rate</th>
			</tr>
			<tr>
				<td>2019</td>
				<td>",
cta_area$total19,
"</td>
				<td>",
round(cta_area$rate19,0),
"</td>
			</tr>
			<tr>
				<td>2020</td>
				<td>",
cta_area$total20,
"</td>
				<td>",
round(cta_area$rate20,0),
"</td>
			</tr>
						<tr>
				<td>2021</td>
				<td>",
cta_area$total21,
"</td>
				<td>",
round(cta_area$rate21,0),
"</td>
			</tr>
								<tr>
				<td>2022</td>
				<td>",
cta_area$total22,
"</td>
				<td>",
round(cta_area$rate22,0),
"</td>
			</tr>
								<tr>
				<td>2023</td>
				<td>",
cta_area$total23,
"</td>
				<td>",
round(cta_area$rate23,0),
"</td>
			</tr>
						<tr>
				<td>2024</td>
				<td>",
cta_area$total24,
"</td>
				<td>",
round(cta_area$rate24,0),
"</td>
			</tr>
						<tr>
				<td>Last 12 months</td>
				<td>",
cta_area$last12mos,
"</td>
				<td>",
round(cta_area$rate_last12,0),
"</td>
			</tr>
						<tr>
				<td>3-year Average</td>
				<td>",
round(cta_area$avg_prior3years,0),
"</td>
				<td>",
round(cta_area$rate_prior3years,0),
"</td>
			</tr>
</table>")

# Creating police beats map for types of crimes
chicago_cta_map <- leaflet(cta_area, options = leafletOptions(zoomControl = FALSE, zoomSnap = 0.5, zoomDelta=0.5)) %>%
  htmlwidgets::onRender("function(el, x) {
L.control.zoom({ position: 'topright' }).addTo(this)
}") %>%
  setView(-87.65, 41.83, zoom = 10.5) %>%
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(color = "white", 
              popup = ctalabel,
              popupOptions = popupOptions(maxWidth ="200", 
                                          minWidth ="200"),
              weight = 2, 
              smoothFactor = 0.5,
              opacity = 0.6, 
              fillOpacity = 0.6,
              fillColor = ~ctapal1(rate_last12),
              group="Rate") %>% 
    addPolygons(color = "white", 
              popup = ctalabel,
              popupOptions = popupOptions(maxWidth ="200", 
                                          minWidth ="200"),
              weight = 2, 
              smoothFactor = 0.5,
              opacity = 0.6, 
              fillOpacity = 0.6,
              fillColor = ~ctapal2(last12mos),
              group="Number") %>% 
    addSearchOSM(options = searchOptions(autoCollapse=FALSE, minLength = 3,zoom=13, position="topleft")) %>%
    onRender("function(el, x) {
        $('input.search-input')[0].placeholder = 'Search street, place or zip code'
        }") %>%
  addLegend(opacity = 0.6,
            values = cta_area$rate_last12, 
            pal = ctapal1,
            position = "bottomleft", 
            title = paste(sep="","<popuptitle>CTA Violent Crime Rate<br><popupchatter>Violent Crimes on CTA/100K people<br>Citywide: ",prettyNum(round(violent_transit$rate_last12,0), big.mark=",")),
            group = "Rate",
            className = "info legend Rate") %>%
  addLegend(opacity = 0.6,
            values = cta_area$last12mos, 
            pal = ctapal2,
            position = "bottomleft", 
            title = paste(sep="","<popuptitle>CTA Violent Crimes<br><popupchatter>Total Violent Crimes on CTA<br>Citywide: ",prettyNum(round(violent_transit$last12mos,0), big.mark=",")),
            group = "Number",
            className = "info legend Number") %>%
  addLayersControl(
    baseGroups = c("Rate","Number"),
    options = layersControlOptions(collapsed = FALSE),
    position = 'bottomright') %>% hideGroup(c("Number")) %>%
   htmlwidgets::onRender("
      function(el, x) {
         var updateLegend = function () {
            var selectedGroup = document.querySelectorAll('input:checked')[0].nextSibling.innerText.substr(1);

            document.querySelectorAll('.legend').forEach(a => a.hidden=true);
            document.querySelectorAll('.legend').forEach(l => {
               if (l.classList.contains(selectedGroup)) l.hidden=false;
            });
         };
         updateLegend();
         this.on('baselayerchange', el => updateLegend());
      }"
   )
chicago_cta_map
```

<br> <br>
<!-- <subhead>CTA Violent Crimes month by month, 2019 through last month</subhead> -->

<iframe title="CTA Violent Crimes month by month since 2019" aria-label="Column Chart" id="datawrapper-chart-dA4Hx" src="https://datawrapper.dwcdn.net/dA4Hx/1/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="400" data-external="1">

</iframe>

```{=html}
<script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(e){if(void 0!==e.data["datawrapper-height"]){var t=document.querySelectorAll("iframe");for(var a in e.data["datawrapper-height"])for(var r=0;r<t.length;r++){if(t[r].contentWindow===e.source)t[r].style.height=e.data["datawrapper-height"][a]+"px"}}}))}();
</script>
```
<br> <br> <!-- <subhead>Violent crime rate comparisons </subhead> -->

<iframe title="2021 violent crime rate comparison in select major cities" aria-label="Bar Chart" id="datawrapper-chart-WJZDu" src="https://datawrapper.dwcdn.net/WJZDu/2/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="436" data-external="1">

</iframe>

```{=html}
<script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(e){if(void 0!==e.data["datawrapper-height"]){var t=document.querySelectorAll("iframe");for(var a in e.data["datawrapper-height"])for(var r=0;r<t.length;r++){if(t[r].contentWindow===e.source)t[r].style.height=e.data["datawrapper-height"][a]+"px"}}}))}();
</script>
```
<br> <br>

```{=html}
<script>!function(){var a,n,i=window||{},s=document||{},t=0;function r(){var e=a.scrollHeight+20;isNaN(e)||e==t||(t=e,isNaN(t)||i.webkit&&i.webkit.messageHandlers&&i.webkit.messageHandlers.espnWebview&&i.webkit.messageHandlers.espnWebview.postMessage({callBack:null,commandName:"ABCNresize",params:{height:t}}))}function o(){var e,a;e=r,a=250,clearTimeout(n),n=setTimeout(e,a)}i.addEventListener("load",function(e){a=s.querySelector(".main-container"),o(),i.addEventListener("resize",o)})}();</script>
```
